\section{Introduction}

Large Language Models (LLMs) such as GPT, ChatGPT, and Vicuna have demonstrated remarkable capabilities in generating fluent, coherent, and contextually relevant text across a wide range of tasks. They leverage vast amounts of textual data and advanced neural architectures to produce responses that often mimic human-level understanding and reasoning \cite{wei2024fewl, ji2023selfreflect}. These models are increasingly being applied in high-stakes domains such as education, healthcare, coding assistance, and knowledge-intensive question-answering systems, where the accuracy and reliability of generated responses are critical.

Despite their impressive performance, LLMs are prone to producing factually incorrect or unsubstantiated information, commonly referred to as hallucinations \cite{wei2024fewl, ji2023selfreflect}. Hallucinations can range from minor factual errors to completely fabricated information, making it challenging to assess model reliability consistently across tasks and domains. This issue is particularly problematic in applications where incorrect responses can lead to misinformation, loss of trust, or potentially harmful consequences \cite{ji2023selfreflect}. The root cause lies in the way LLMs are trained: they optimize for linguistic fluency and contextual plausibility rather than factual correctness, which can result in responses that appear convincing but are unreliable.

Detecting hallucinations in LLM outputs has therefore become a critical area of research. Methods such as \textbf{ChainPoll} \cite{friel2023chainpoll} propose high-efficacy metrics and realistic benchmarks to evaluate hallucination detection. ChainPoll introduces RealHall, a suite of challenging datasets designed to reflect practical LLM usage, and demonstrates superior accuracy compared to prior metrics such as SelfCheckGPT, GPTScore, and G-Eval. Detection frameworks like these provide insights into the reliability of generated outputs, highlight limitations in model behavior, and guide strategies for model improvement.

Beyond detection, measuring and benchmarking hallucinations remains challenging, especially when human-written gold-standard answers are unavailable. Recent approaches, including \textbf{FEWL} \cite{wei2024fewl}, leverage off-the-shelf LLMs as reference models to continuously score hallucinations. FEWL estimates each reference model’s expertise by evaluating agreement with incorrect answers and penalizing superficial knowledge, allowing reliable hallucination measurement without costly human annotations. Such approaches are crucial for creating scalable, cost-effective benchmarks that more accurately reflect real-world LLM behavior.

Several studies have also explored mitigation strategies to reduce hallucinations in LLMs. Self-reflection methods \cite{ji2023selfreflect} iteratively generate background knowledge, evaluate factuality, and refine responses to improve answer correctness. In addition, approaches such as in-context learning and supervised fine-tuning can leverage metrics like FEWL to guide LLMs toward generating more factual outputs \cite{wei2024fewl}. Together, these detection, measurement, and mitigation techniques enhance the reliability of LLMs in diverse domains, including medical question-answering, coding assistance, and other knowledge-intensive tasks.

This survey aims to provide a structured overview of hallucination research in LLMs, focusing on three key areas: detection, measurement, and mitigation. Our anticipated contributions include: (1) summarizing recent advances in hallucination detection and measurement frameworks such as FEWL and ChainPoll, (2) highlighting mitigation strategies including self-reflection and fine-tuning, and (3) providing practical insights for designing more trustworthy and reliable LLM systems suitable for deployment in real-world applications.